---
title: "BertSRC_data"
output: html_document
date: "2023-07-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document combines the test/training/validation data from BertSRC (https://github.com/ssr01357/BertSRC/tree/master/data).
The combined data is filtered for gene/protein interactions (gene-gene, gene-protein, protein-protein).
The BertSRC data only contains the sentence from which the interaction was sourced. We wanted to look at the entire abstract.
Using these interactions and easyPubMed, pulled the entire abstract for each interaction PMID and combined the original data with their 
matched abstracts. The function is called extractAbstracts(), and expects a dataframe as input (the combined and filtered data mentioned above). 
This input dataframe must have the structure of the BertSRC data: 17 columns, with PMID in the 1st column. 

## Read in the data:
```{r}
test_data = read.csv('./test.csv', stringsAsFactors = FALSE)
```

```{r}
train_data = read.csv('./train.csv', stringsAsFactors = FALSE)
```

```{r}
validation_data = read.csv('./validation.csv', stringsAsFactors = FALSE)
```

```{r}
library(dplyr)
library(tidyverse)
```

Combine the data into one data frame by column. 
```{r}
data_all <- list.files(path = "../BertSRC_data/",  # Identify all CSV files
                       pattern = "*.csv", full.names = TRUE) %>% 
  lapply(read.csv) %>%                              # Store all files in list
  bind_rows                                         # Combine data sets into one data set 
data_all    

```

Filter data_all by gene-gene and protein-protein interactions.
column 4 and column 9 must be both gene or both protein
Update 9/11/23: column 4 and 9 can be both/either gene/protein - all combos
```{r}
entities = c("GENE", "PROTEIN")

#gg_data = data_all %>% filter((X4 == 'GENE') & (X9 == "GENE"))
#pp_data = data_all %>% filter((X4 == "PROTEIN") & (X9 == "PROTEIN"))
gg_pp_data = data_all %>% filter( ((X4 == "GENE") & (X9 == "GENE")) | ((X4 == "PROTEIN") & (X9 == "PROTEIN")) )
gene_prot_all_data = data_all %>% filter( ((X4 == "GENE") & (X9 == "GENE")) | ((X4 == "PROTEIN") & (X9 == "PROTEIN")) | 
                                   ((X4 == "GENE") & (X9 == "PROTEIN")) | ((X4 == "PROTEIN") & (X9 == "GENE")) )
#gene_prot_all_data_1 = data_all %>% filter( ((X4 == ("GENE" | "PROTEIN")) | (X9 == ("GENE" | "PROTEIN")) ) )
gene_prot_all_data_1 = data_all %>% filter((X4 %in% entities) & (X9 %in% entities))

```


```{r}
install.packages("easyPubMed")
```


```{r}
library(easyPubMed)
```

```{r}
easyPubMed::article_to_df(gg_pp_data$X0[1])
```


https://www.data-pulse.com/projects/Rlibs/vignettes/easyPubMed_02_advanced_tutorial.html
Demo 1:
```{r}
pmids = as.list(gg_pp_data$X0)

example_query = get_pubmed_ids(pmids[1])
example_xml = fetch_pubmed_data(example_query)
example_xml_list = articles_to_list(example_xml)
example_df = do.call(rbind, lapply(example_xml_list, article_to_df, max_chars = -1, getAuthors = FALSE))


```


Use easyPubMed and Demo 1 example on my data:
```{r}
# make a dataframe to hold 2 columns: pmids and title+abstract
pubmed_abstracts = data.frame(matrix(ncol = 2, nrow = length(pmids)))
col_headers = c("pmid", "title_abstract")
colnames(pubmed_abstracts) = col_headers
# populate pmid column
pubmed_abstracts$pmid = pmids
```


```{r}
# do a for loop for all pmid's of interest
for(i in 1:length(pmids)) {
  paper = pmids[i]
  query_pubmed_server = get_pubmed_ids(paper)
  pubmed_xml = fetch_pubmed_data(query_pubmed_server)
  xml_list = articles_to_list(pubmed_xml)
  pubmed_record = do.call(rbind, lapply(xml_list,
                                        article_to_df,
                                        max_chars = -1,
                                        getAuthors = FALSE))
  # populate dataframe with extracted information (title and abstract)
  pubmed_abstracts$title_abstract[i] = paste(pubmed_record$title, pubmed_record$abstract)
}
  
```


Update combined data with extracted pubmed titles and abstracts
Replace column 14 with pubmed_abstracts$title_abstract

```{r}

for(i in 1:length(pmids)) {
  gg_pp_data$X14[i] = pubmed_abstracts$title_abstract[i]
}

```

Make column names more human readable:
```{r}
gg_pp_col_headers = c("PMID", "Sentence_ID", "Left_Context_Entity1", "Entity1", "Entity1_Type", "Right_Context_Entity1",
                      "Biological_Context", "Left_Context_Entity2", "Entity2", "Entity2_Type", "Right_Context_Entity2",
                      "Directional_Relation", "Activity_Relation", "Relation_Verb", "Title_Abstract", 
                      "Semantic_Relation_1", "Semantic_Relation_2")
```

assign new column names to the data table
```{r}
colnames(gg_pp_data) = gg_pp_col_headers
```


```{r}
orig_sentence = data_all %>% filter( ((X4 == "GENE") & (X9 == "GENE")) | ((X4 == "PROTEIN") & (X9 == "PROTEIN")) )
# original sentence = orig_sentence$X14
```


```{r}
gg_pp_data$Sentence = orig_sentence$X14
```


Write out the data to a tab-delimited txt file.
```{r}
write.table(gg_pp_data, file = "BertSRC_data.txt", sep = "\t", col.names = TRUE, row.names = FALSE)
```

-----------------------------
```{r}
library(easyPubMed)
```

```{r}
pmids = as.list(gene_prot_all_data_1$X0)

# make a dataframe to hold 2 columns: pmids and title+abstract
pubmed_abstracts = data.frame(matrix(ncol = 2, nrow = length(pmids)))
col_headers = c("pmid", "title_abstract")
colnames(pubmed_abstracts) = col_headers
# populate pmid column
pubmed_abstracts$pmid = pmids

```


```{r}
# do a for loop for all pmid's of interest
for(i in 1:length(pmids)) {
  paper = pmids[i]
  query_pubmed_server = get_pubmed_ids(paper)
  pubmed_xml = fetch_pubmed_data(query_pubmed_server)
  xml_list = articles_to_list(pubmed_xml)
  pubmed_record = do.call(rbind, lapply(xml_list,
                                        article_to_df,
                                        max_chars = -1,
                                        getAuthors = FALSE))
  # populate dataframe with extracted information (title and abstract)
  pubmed_abstracts$title_abstract[i] = paste(pubmed_record$title, pubmed_record$abstract)
}
```


```{r}
# Update combined data with extracted pubmed titles and abstracts
# Add title&abstract column @ X17 with pubmed_abstracts$title_abstract
for(i in 1:length(pmids)) {
  gene_prot_all_data_1$X17[i] = pubmed_abstracts$title_abstract[i]
}

# make column headers
gg_pp_col_headers = c("PMID", "Sentence_ID", "Left_Context_Entity1", "Entity1", "Entity1_Type", "Right_Context_Entity1",
                      "Biological_Context", "Left_Context_Entity2", "Entity2", "Entity2_Type", "Right_Context_Entity2",
                      "Directional_Relation", "Activity_Relation", "Relation_Verb", "Sentence", 
                      "Semantic_Relation_1", "Semantic_Relation_2", "Title_Abstract")
colnames(gene_prot_all_data_1) = gg_pp_col_headers

#orig_sentence = data_all %>% filter( ((X4 == "GENE") & (X9 == "GENE")) | ((X4 == "PROTEIN") & (X9 == "PROTEIN")) )
#gg_pp_data$Sentence = orig_sentence$X14

#write.table(gg_pp_data, file = "BertSRC_data.txt", sep = "\t", col.names = TRUE, row.names = FALSE)
write.table(gene_prot_all_data_1, file = "BertSRC_data_augmented.txt", sep = "\t", col.names = TRUE, row.names = FALSE)
```




```{r}
# extractAbstracts takes a filtered dataframe as input
# the filtered dataframe is in the format from the BertSRC data
# filtered for specific entities in column X4 and X9
# function uses PMID and package easyPubMed to extract abstracts
# and adds them to output dataframe, which gets written out to a tab-delimited 
# text file
extractAbstracts = function(bertsrc_data) {
  pmids = as.list(bertsrc_data$X0)
  # make a dataframe to hold 2 columns: pmids and title+abstract
  pubmed_abstracts = data.frame(matrix(ncol = 2, nrow = length(pmids)))
  col_headers = c("pmid", "title_abstract")
  colnames(pubmed_abstracts) = col_headers
  # populate pmid column
  pubmed_abstracts$pmid = pmids
  
  # do a for loop for all pmid's of interest
  for(i in 1:length(pmids)) {
    paper = pmids[i]
    query_pubmed_server = get_pubmed_ids(paper)
    pubmed_xml = fetch_pubmed_data(query_pubmed_server)
    xml_list = articles_to_list(pubmed_xml)
    pubmed_record = do.call(rbind, lapply(xml_list,
                                          article_to_df,
                                          max_chars = -1,
                                          getAuthors = FALSE))
    # populate dataframe with extracted information (title and abstract)
    pubmed_abstracts$title_abstract[i] = paste(pubmed_record$title, pubmed_record$abstract)
  }
    
  # Update combined data with extracted pubmed titles and abstracts
  # Add title&abstract column @ X17 with pubmed_abstracts$title_abstract
  for(i in 1:length(pmids)) {
    bertsrc_data$X17[i] = pubmed_abstracts$title_abstract[i]
  }
  
  # make column headers
  gg_pp_col_headers = c("PMID", "Sentence_ID", "Left_Context_Entity1", "Entity1", "Entity1_Type", "Right_Context_Entity1",
                      "Biological_Context", "Left_Context_Entity2", "Entity2", "Entity2_Type", "Right_Context_Entity2",
                      "Directional_Relation", "Activity_Relation", "Relation_Verb", "Sentence", 
                      "Semantic_Relation_1", "Semantic_Relation_2", "Title_Abstract")
  colnames(bertsrc_data) = gg_pp_col_headers
  
  return(bertsrc_data)
}
```


```{r}
bertsrc_abstract_data = extractAbstracts(gene_prot_all_data)
```


