---
title: "intact_xml_parser_scratchpad_2"
output: html_document
date: "2023-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Other scratchpad got too cluttered
fresh page

```{r}
library(RpsiXML)
#devtools::install_github("bioc/RpsiXML")
library(tidyverse)
library(doParallel)
library(foreach)
```


```{r}
make_unique_ID <- function(PMID, term1, term2) {
  paste0(c(PMID, sort(c(term1, term2))), collapse = " ")
}

deDupe = function(interaction_df) {
  interaction_df_id = interaction_df %>%
    mutate(ID = pmap(list(PMID = PMID, term1 = EntityA, term2 = EntityB), make_unique_ID) %>% unlist())
  
  deduped = interaction_df_id %>% distinct(ID, .keep_all = TRUE)
  return(deduped)
}

```



```{r}
pull_interactions = function(taxonID) {
  setwd("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI")
  if (dir.exists(paths = "./IntAct/psi25/species/") == FALSE) {
    print("Cannot find appropriate species files. Download all.zip from IntAct and unzip psi25/species/.")
  }
  
  # all of IntAct:
  filenames = list.files("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/IntAct/psi25/species/", 
                         pattern ="*.xml", full.names = TRUE)
  
  # THIS IS YEAST TEST. 
  #filenames = list.files("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/IntAct/yeast_test", 
  #                       pattern = "*.xml", full.names = TRUE)
  
  species_interactions = data.frame()
  data_headers = c("PMID", "EntityA", "EntityB", "SpeciesA", "TaxonID_A", 
                   "SpeciesB", "TaxonID_B")
  
  unfiltered_interactions = 0 #keeping count of interactions being thrown out
  total_interactions = 0
  
  # parallelize pulling interactions:
  # don't use all the cores on machine since that will break things
  numcores <- parallel::detectCores()/2 
  cl <- parallel::makeCluster(numcores) 
  clusterEvalQ(cl = cl, library(magrittr)) 
  clusterEvalQ(cl = cl, library(dplyr)) # clusterExport(cl = cl, "combos_num") library(foreach) library(doParallel) 
  registerDoParallel(cl)
  
  # results is the dataframe housing all the ppi data
  # RESULTS = PARALLELIZED MODE
  results = foreach(i = 1:length(filenames), .combine = "rbind", .packages = 'RpsiXML') %dopar% {
  # FOR.. = FOR-LOOP MODE
  #for(i in 1:length(filenames)) {
    # current_interaction is the S4 object made by 1 file
    current_interaction = RpsiXML::parsePsimi25Interaction(filenames[i], INTACT.PSIMI25, verbose=TRUE) 
    record = data.frame(matrix(nrow = length(RpsiXML::interactions(current_interaction)), ncol = length(data_headers)))
    
    # within each file, there are n interactions:
    for(j in 1:length(RpsiXML::interactions(current_interaction))) {
      pmid = as.numeric(current_interaction@interactions[[j]]@expPubMed[[1]])
      bait = na.omit(current_interaction@interactions[[j]]@baitUniProt)
      prey_all = na.omit(current_interaction@interactions[[j]]@preyUniProt)
      #participants_all = current_interaction@interactions[[j]]@participant
      
      # check bait first:
      # if length of bait != 1, next
      if(length(bait) != 1){
        unfiltered_interactions = unfiltered_interactions + 1
        total_interactions = total_interactions + 1
        next
      # if bait is NA, next
      } else if (anyNA(bait)) {
        unfiltered_interactions = unfiltered_interactions + 1
        total_interactions = total_interactions + 1
        next
          
      # check prey next:
        # if there's less than one prey entry, next
      } else if (length(prey_all) < 1) {
        unfiltered_interactions = unfiltered_interactions + 1
        total_interactions = total_interactions + 1
        next
        
        # checking if there's exactly 1 prey:
      } else if (length(prey_all) == 1) {
          # if our single prey is NA, next
          if (anyNA(prey_all)) { 
            unfiltered_interactions = unfiltered_interactions + 1
            total_interactions = total_interactions + 1
            next
          
          # if prey == bait, next
          } else if (bait == prey_all) {
            unfiltered_interactions = unfiltered_interactions + 1
            total_interactions = total_interactions + 1
            next
          } else {
            # at this point, we know:
            # bait is valid protein
            # prey is also one valid protein that isn't the same as bait
            
            speciesA = lapply(current_interaction@interactors[bait], function(x) x@organismName)[[1]]
            speciesA_taxon = lapply(current_interaction@interactors[bait], function(x) x@taxId)[[1]]
            speciesB = lapply(current_interaction@interactors[prey_all], function(x) x@organismName)[[1]]
            speciesB_taxon = lapply(current_interaction@interactors[prey_all], function(x) x@taxId)[[1]]
            
            if((speciesA_taxon == taxonID) & (speciesB_taxon == taxonID)) {
              record_list = c(pmid, bait, prey_all, speciesA, speciesA_taxon, speciesB, speciesB_taxon)
              record[j, ] = record_list
              total_interactions = total_interactions + 1
            } else {
              unfiltered_interactions = unfiltered_interactions + 1
              total_interactions = total_interactions + 1
            }
          }
          
          # more than 1 prey:
        } else if(length(prey_all) > 1) {
          
          for(k in 1:length(prey_all)) {
            prey = prey_all[k]
            speciesA = lapply(current_interaction@interactors[bait], function(x) x@organismName)[[1]]
            speciesA_taxon = lapply(current_interaction@interactors[bait], function(x) x@taxId)[[1]]
            speciesB = lapply(current_interaction@interactors[prey], function(x) x@organismName)[[1]]
            speciesB_taxon = lapply(current_interaction@interactors[prey], function(x) x@taxId)[[1]]
            
            if((speciesA_taxon == taxonID) & (speciesB_taxon == taxonID)) {
              record_list = c(pmid, bait, prey, speciesA, speciesA_taxon, speciesB, speciesB_taxon)
              record[j, ] = record_list
              total_interactions = total_interactions + 1
            } else {
              unfiltered_interactions = unfiltered_interactions + 1
              total_interactions = total_interactions + 1
            }
            
          }
          
        }
      } # end of inner for-loop
      species_interactions = rbind(species_interactions, record)
    } # end of outer for-loop
  stopCluster(cl)
  
  # these two lines for parallel run
  colnames(results) = data_headers
  species_interactions = na.omit(results)
  
  # these two lines for for-looping
  #colnames(species_interactions) = data_headers
  #species_interactions = na.omit(species_interactions)
  
  deduped_interactions = deDupe(species_interactions)
  num_dupe = nrow(species_interactions) - nrow(deduped_interactions)
  
  # some summary stats: (only for for-looping)
  #outfile_name = paste0(taxonID, "_summary_stats.txt")
  #write_lines(
  #  c(paste0("organism: ", taxonID),
  #    paste0("invalid interactions tally: ", unfiltered_interactions),
  #    paste0("total number of interactions: ", total_interactions),
  #    paste0("duplicates tally: ", num_dupe),
  #    paste0("percentage of interactions thrown out: ", (unfiltered_interactions + num_dupe)/total_interactions)), 
  #  file = outfile_name
  #)
  
  data_outfile_name = paste0(taxonID, "_interactions_test.csv")
  write.table(deduped_interactions, 
              file = data_outfile_name,
              quote = FALSE,
              sep = '\t',
              row.names = FALSE)
  
  return(deduped_interactions)
} # end of function


# PARALLEL MODE: CHANGE ME WHEN CODE IS UPDATED
```


```{r}
start.time = Sys.time()
pull_interactions(83333) #E.coli
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```
E.coli runtime on parallelized code = 41.15352 minutes. 
Try again on nested for-loop to gather some metadata? - nope, kill, took days without results. 




```{r}
start.time = Sys.time()
pull_interactions(559292) #yeast
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

yeast_test had 593 entries 
update the "next"
deduper doesn't work still

now it's 404 entries, deduper is definitely working
not entirely sure if the for-loop for multiple prey are working

with more yeast files, it's extended to 3,690 entries
but I can't troubleshoot prey_all - multiple entries on parallel mode

back to for-loop mode, with fewer files
multiple prey loop IS triggered, 590 entries on yeast files 1 - 4
try the same in parallel mode - trying to see if same entries results
same number of results, 590. 

pretty sure with deduper working and multiple prey working, try for ecoli again


```{r}
ecoli_interactions = read.csv("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/83333_interactions.csv")
ecoli_interactions = ecoli_interactions %>% mutate(ID = pmap(list(PMID = PMID, term1 = EntityA, term2 = EntityB), make_ID) %>% unlist())
make_ID <- function(PMID, term1, term2) {
    paste0(c(PMID, sort(c(term1, term2))), collapse = " ")
}
ecoli_interactions_dedupe = ecoli_interactions %>% distinct(ID, .keep_all = TRUE)
make_interaction = function(term1, term2) {paste0(sort(c(term1, term2)), collapse = " ")}
ecoli_interactions_dedupe = ecoli_interactions_dedupe %>% mutate(Interaction = pmap(list(term1 = EntityA, term2 = EntityB), make_interaction) %>% unlist())

ecoli_interactions_dedupe = ecoli_interactions_dedupe %>% group_by(Interaction) %>% mutate(Abstract_Count = n())
ecoli_interactions_dedupe = ecoli_interactions_dedupe %>% group_by(PMID) %>% mutate(pmid_count = n())

write.table(ecoli_interactions_dedupe, file = "83333_interactions_deduped.txt", quote = FALSE, sep = '\t', row.names = FALSE)
```


