---
title: "uniprotR_interactions"
output: html_document
date: "2024-01-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(UniprotR)
library(tidyverse)
library(data.table)
```


```{r}
lysA = GetProteinInteractions("P00861")
```

```{r}
ecoli_83333 = fread("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/litportal_ecoli/uniprotkb_proteome_UP000000625_2023_11_27.txt")
```


uniprotkb_taxonomy_id_562_2024_01_25.tsv was downloaded from UniProt's website, using the UniProtKB searchbar, taxonomy ID: 562. The download file was as a zipped .tsv file, and unzipped in terminal using gunzip [filename].

```{r}
ecoli_562 = fread("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/uniprotkb_taxonomy_id_562_2024_01_25.tsv")
```

```{r}
reviewed_562 = ecoli_562 %>% filter(Reviewed == "reviewed")
table(ecoli_562$Reviewed)
```

Build list of proteins for 562:
```{r}
unique_562 = data.frame(
  Entry = reviewed_562$Entry,
  #Entry_Name = reviewed_562$`Entry Name`,
  Protein_Names = reviewed_562$`Protein names`
  #Organism = reviewed_562$Organism
)
```


Make file for list of proteins for 562:
```{r}
fwrite(unique_562, file = "/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/562_ecoli_proteins.csv", sep = ",")
```

Make dataframe of Protein A and B
```{r}
interactions_562 = data.frame(
  Protein_A = reviewed_562$Entry,
  Protein_B = reviewed_562$`Interacts with`
)
```

Isolate the entries with interactions: the entries where "Protein_B" is populated
```{r}
interactions_562 = interactions_562[!(is.na(interactions_562$Protein_B) | interactions_562$Protein_B==""),]
```

Of the interactions, convert entries with multiple Protein_B into binary interactions
for example: 
ProteinA = A, ProteinB = B, C, D, E 
==> 
A, B
A, C
A, D
A, E
```{r}
binary_interactions_562 = data.frame(matrix(ncol=2))
colnames(binary_interactions_562) = c("Protein_A", "Protein_B")
bin_int_count = 0
mult_int_count = 0
for(i in seq(dim(interactions_562)[1])) {
  interacts_with = strsplit(interactions_562$Protein_B[i], "; ") %>% unlist()
  if (length(interacts_with) == 1){
    bin_int_count = bin_int_count + 1
    proteinA = interactions_562$Protein_A[i]
    proteinB = interactions_562$Protein_B[i]
    binary_interaction = c(proteinA, proteinB)
    binary_interactions_562 = rbind(binary_interactions_562, binary_interaction)
  } else { # more than 1 Protein_B
    mult_int_count = mult_int_count + 1
    for(j in seq(length(interacts_with))){
      proteinA = interactions_562$Protein_A[i]
      proteinB = interacts_with[j]
      binary_interaction = c(proteinA, proteinB)
      binary_interactions_562 = rbind(binary_interactions_562, binary_interaction)
    }
  }
}
binary_interactions_562 = na.omit(binary_interactions_562)
```

Filter proteins that interact with itself: Protein A == Protein B
```{r}
binary_interactions_562_dedupe = binary_interactions_562 %>% filter(Protein_A != Protein_B)
```

Filter duplicate interactions:
for example, 
Protein A : Protein B 
Protein B : Protein A 
Keep just one

make_interaction_ID makes a unique interaction ID by sorting each term alphabetically
```{r}
make_interaction_ID <- function(term1, term2) {
  paste0(sort(c(term1, term2)), collapse = " ")
}
```

Make unique identifier for each binary interaction:
```{r}
binary_interactions_562_dedupe_int = binary_interactions_562_dedupe %>% mutate(
  ID = pmap(list(term1 = Protein_A, term2 = Protein_B), make_interaction_ID) %>% unlist()
)
```

Keep interactions that are unique 
```{r}
binary_interactions_562_dedupe_int_dist = binary_interactions_562_dedupe_int %>% distinct(ID, .keep_all = TRUE)
```


```{r}
getPMID_562 = function(proteinA, proteinB) {
  interacts_with = ecoli_562[(ecoli_562$Entry == proteinA),]$`Interacts with` %>% strsplit("; ") %>% unlist()
  if(proteinB %in% interacts_with) {
    pmids = ecoli_562[(ecoli_562$Entry == proteinA),]$`PubMed ID`
  } else {
    pmids = NA
  }
  return(pmids)
}
```


```{r}
ecoli_562_interactions_table_pubs = binary_interactions_562_dedupe_int_dist %>% mutate(
  PMIDs = map2_chr(Protein_A, Protein_B, getPMID_562)
)
```


```{r}
ecoli_562_papers = ecoli_562_interactions_table_pubs$PMIDs %>% strsplit("; ") %>% compact %>% unlist() %>% unique()
```

```{r}
fwrite(as.list(ecoli_562_papers), file = "/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/DATABASE_PIPELINE/benchmark3/ecoli_562_papers.csv", sep = ",")
```



test case:
```{r}
test_interactions = head(interactions_562, 45)
```

```{r}
test_bin_int = data.frame(matrix(ncol=2))
colnames(test_bin_int) = c("Protein_A", "Protein_B")

for(i in seq(dim(test_interactions)[1])){
  test_int_with = strsplit(test_interactions$Protein_B[i], "; ") %>% unlist()
  if (sum(grepl("\\s+", test_int_with)) > 0) {test_int_with = str_extract(test_int_with, "(?<=\\[).*(?=\\])")}
  if (length(test_int_with) == 1) {
    proteinA = test_interactions$Protein_A[i]
    proteinB = test_int_with
    bin_int = c(proteinA, proteinB)
    test_bin_int = rbind(test_bin_int, bin_int)
  } else {
    for(j in seq(length(test_int_with))) {
      proteinA = test_interactions$Protein_A[i]
      proteinB = test_int_with[j]
      bin_int = c(proteinA, proteinB)
      test_bin_int = rbind(test_bin_int, bin_int)
    }
  }
}
test_bin_int = na.omit(test_bin_int)
```


```{r}
test_bin_int_dedupe = test_bin_int %>% filter(Protein_A != Protein_B)
```



Continue building dataframe with Species and publication information:

```{r}
binary_interactions_562 = binary_interactions_562 %>% mutate(
  Species_A = map(Protein_A, getOrganism),
  Species_B = map(Protein_B, getOrganism),
  Number_Publications = map2_dbl(Protein_A, Protein_B, getBinIntPubs)
)
```
slow!
url error? ---> need to clean data more

```{r}
binary_interactions_562 = binary_interactions_562 %>% mutate(
  Species_A = GetProteinAnnontate(Protein_A, "organism_name"),
  Species_B = GetProteinAnnontate(Protein_B, "organism_name"),
  Number_Publications = map2_dbl(Protein_A, Protein_B, getBinIntPubs)
)
```


function for finding the number of publications per binary interaction
```{r}
getBinIntPubs = function(proteinA, proteinB) {
  A = ecoli_562 %>% filter(Entry == proteinA)
  pubs_A = A$`PubMed ID` %>% strsplit("; ") %>% unlist()
  B = ecoli_562 %>% filter(Entry == proteinB)
  pubs_B = B$`PubMed ID` %>% strsplit("; ") %>% unlist()
  #print(c("proteinA: ", proteinA))
  #print(c("pubs_A: ", pubs_A))
  #print(c("proteinB: ", proteinB))
  #print(c("pubs_B: ", pubs_B))
  
  interaction_pubs = intersect(pubs_A, pubs_B)
  #print(c("interaction_pubs: ", interaction_pubs))
  return(length(interaction_pubs))
}
```



```{r}
getBinIntPubs0 = function(proteinA, proteinB) {
  pubs_A = GetProteinAnnontate(proteinA, "lit_pubmed_id") %>% strsplit("; ") %>% unlist()
  pubs_B = GetProteinAnnontate(proteinB, "lit_pubmed_id") %>% strsplit("; ") %>% unlist()
  #print(c("proteinA: ", proteinA))
  #print(c("pubs_A: ", pubs_A))
  #print(c("proteinB: ", proteinB))
  #print(c("pubs_B: ", pubs_B))
  
  interaction_pubs = intersect(pubs_A, pubs_B)
  #print(c("interaction_pubs: ", interaction_pubs))
  return(length(interaction_pubs))
}
```



```{r}
getOrganism = function(protein) {
  protein_info = ecoli_562 %>% filter(Entry == protein)
  #print(protein_info)
  organism_name = protein_info$Organism 
  #print(organism_name)
  return(organism_name)
}
```



```{r}
bin_int_test0 = test_bin_int_dedupe %>% mutate(
  Species_A = GetProteinAnnontate(Protein_A, "organism_name"),
  Species_B = GetProteinAnnontate(Protein_B, "organism_name"),
  Number_Publications = map2_dbl(Protein_A, Protein_B, getBinIntPubs0)
)
```
2:38p - 2:44p, 6 minutes for 88 entries, 88*2 proteins, 88*2*3 GetProteinAnnontate() queries

parallelize:
3:41p
```{r}
library(doParallel)
library(foreach)
# parallelize pulling interactions:
# don't use all the cores on machine since that will break things
numcores <- parallel::detectCores()/2 
cl <- parallel::makeCluster(numcores) 
clusterEvalQ(cl = cl, library(magrittr)) 
clusterEvalQ(cl = cl, library(dplyr)) # clusterExport(cl = cl, "combos_num") library(foreach) library(doParallel) 
registerDoParallel(cl)

start.time = Sys.time()
results = foreach(i = 1:dim(test_bin_int_dedupe)[1], .combine = "cbind", .packages = c('UniprotR', 'tidyverse')) %dopar% {
  bin_int_test1 = test_bin_int_dedupe %>% mutate(
  Species_A = GetProteinAnnontate(Protein_A, "organism_name"),
  Species_B = GetProteinAnnontate(Protein_B, "organism_name"),
  Number_Publications = map2_dbl(Protein_A, Protein_B, getBinIntPubs0) )
} 
stopCluster(cl)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```


furr 
```{r}

```



```{r}
bin_int_test2 = test_bin_int %>% mutate(
  Species_A = GetProteinAnnontate(Protein_A, "organism_name"),
  Species_B = GetProteinAnnontate(Protein_B, "organism_name"),
  Number_Publications = map2_dbl(Protein_A, Protein_B, getBinIntPubs)
)
```


```{r}
bin_int_test3 = test_bin_int %>% mutate(
  Species_A = map(Protein_A, getOrganism),
  Species_B = map(Protein_B, getOrganism),
  Number_Publications = map2_dbl(Protein_A, Protein_B, getBinIntPubs)
)
```


```{r}
fwrite(binary_interactions_562, file = "/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/562_ecoli_binary_interactions.tsv", sep = "\t")
```


