---
title: "Uniprot_package_ppi_sjkim"
output: html_document
date: "2023-06-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## UniProt

Downloaded all the Taxon ID's from UniProt's website: https://www.uniprot.org/taxonomy/?query=*
Columns: Taxon Id, Common name, Scientific name, Synonyms, Lineage, Links

```{bash}
cd /Users/kimg044/OneDrive\ -\ PNNL/Documents/PPI
wc -l uniprot-compressed_true_download_true_fields_id_2Ccommon_name_2Cscie-2023.06.12-20.50.17.85.tsv
```

```{r}
taxonIDs = data.table::fread("/Users/kimg044/OneDrive\ -\ PNNL/Documents/PPI/uniprot-compressed_true_download_true_fields_id_2Ccommon_name_2Cscie-2023.06.12-20.50.17.85.tsv")
```

```{r}
# number of taxonIDs
nrow(taxonIDs)
colnames(taxonIDs)

```

```{r}
taxonIDs %>% 
  filter(`Taxon Id` %in% 9506:9706)
```

some taxonIDs have maybe one protein 
2.5 million taxon ID's, hierarchical, and can explore more from the NCBI taxonomy browser: https://www.ncbi.nlm.nih.gov/taxonomy
we'd need relevant taxonomy ID 

```{bash}
head uniprot-compressed_true_download_true_fields_id_2Ccommon_name_2Cscie-2023.06.12-20.50.17.85.tsv
```
Yeast Example:
Yeast taxonomy ID: 559292
https://www.uniprot.org/uniprotkb?facets=reviewed:true&query=(taxonomy_id:559292)
- customize columns to include pathway and interactions
- tsv download

learn how to do API calls
benchmark off of E.coli


## UniProtR
cran package: https://cran.r-project.org/web/packages/UniprotR/index.html
```{r}
install.packages("UniprotR")
```

documentation: https://cran.r-project.org/web/packages/UniprotR/UniprotR.pdf
GetProteinInteractions()
GetproteinNetwork()
GetproteinNetwork_all()
All these functions require a proteins accession list as well as a taxon ID 
These functions also just map to stringDB

https://github.com/Proteomicslab57357/UniprotR/tree/master/R

There isn't a function for all related proteins in a taxonID call 


