---
title: "build_interaction_table"
output: html_document
date: "2024-02-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(data.table)
```

Input:
UniProt data for species of interest
Download all proteins for a species of interest from uniprot.org

Output: 
1: table of binary interactions: 
Protein A | Protein B | Species A | Species B | 
2: list of unique proteins


```{r}
covid_2697049 = fread("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/DATABASE_PIPELINE/benchmark2/uniprotkb_2697049_2024_02_05.tsv")
```

Confirm list of proteins are unique:
```{r}
length(unique(covid_2697049$Entry)) == length(covid_2697049$Entry)
```

```{r}
covid_2697049$Reviewed %>% table()
```
Since there are only 17 reviewed cases in this data table, continue with using all unreviewed and reviewed cases.


Build data.frame of unique proteins:
```{r}
unique_2697049 = data.frame(
  Entry = covid_2697049$Entry,
  Protein_Names = covid_2697049$`Protein names`
)
```

Write data.frame of unique proteins to file:

```{r}
fwrite(unique_2697049, file = "/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/DATABASE_PIPELINE/benchmark2/2697049_proteins.csv", sep = ",")
```

### Build Interactions Table:

Isolate Protein A and Protein B from data:
```{r}
interactions_2697049 = data.frame(
  Protein_A = covid_2697049$Entry,
  Protein_B = covid_2697049$`Interacts with`
)
```


Filter to just the ones where "protein B/Interacts with" column is populated:
```{r}
interactions_2697049 = interactions_2697049[!(is.na(interactions_2697049$Protein_B) | interactions_2697049$Protein_B == ""),]
```


Of the interactions, convert entries with multiple Protein_B into binary interactions
for example: 
ProteinA = A, ProteinB = B, C, D, E 
==> 
A, B
A, C
A, D
A, E

```{r}
binary_interactions_covid = data.frame(matrix(ncol=2))
colnames(binary_interactions_covid) = c("Protein_A", "Protein_B")
bin_int_count = 0
mult_int_count = 0
for(i in seq(dim(interactions_2697049)[1])) {
  interacts_with = strsplit(interactions_2697049$Protein_B[i], "; ") %>% unlist()
  if (length(interacts_with) == 1){
    bin_int_count = bin_int_count + 1
    proteinA = interactions_2697049$Protein_A[i]
    proteinB = interactions_2697049$Protein_B[i]
    binary_interaction = c(proteinA, proteinB)
    binary_interactions_covid = rbind(binary_interactions_covid, binary_interaction)
  } else { # more than 1 Protein_B
    mult_int_count = mult_int_count + 1
    for(j in seq(length(interacts_with))){
      proteinA = interactions_2697049$Protein_A[i]
      proteinB = interacts_with[j]
      binary_interaction = c(proteinA, proteinB)
      binary_interactions_covid = rbind(binary_interactions_covid, binary_interaction)
    }
  }
}
binary_interactions_covid = na.omit(binary_interactions_covid)

```
340 binary interactions in covid, so far

Filter proteins that interact with itself: Protein A == Protein B
```{r}
binary_interactions_covid_dedupe = binary_interactions_covid %>% filter(Protein_A != Protein_B)
```
333 binary interactions in covid, so far

Filter duplicate interactions:
for example, 
Protein A : Protein B 
Protein B : Protein A 
Keep just one

make_interaction_ID makes a unique interaction ID by sorting each term alphabetically
```{r}
make_interaction_ID <- function(term1, term2) {
  paste0(sort(c(term1, term2)), collapse = " ")
}
```

Make unique identifier for each binary interaction:
```{r}
binary_interactions_covid_dedupe_int = binary_interactions_covid_dedupe %>% mutate(
  ID = pmap(list(term1 = Protein_A, term2 = Protein_B), make_interaction_ID) %>% unlist()
)
```

Keep interactions that are unique 
```{r}
binary_interactions_covid_dedupe_int_dist = binary_interactions_covid_dedupe_int %>% distinct(ID, .keep_all = TRUE)
```
300 unique binary interactions in covid


Fill in Species information:
```{r}
getOrganism = function(protein) {
  protein_info = covid_2697049 %>% filter(Entry == protein)
  
  if(length(protein_info$Organism)>0) {
    organism_name = protein_info$Organism
  } else {
    organism_name = "Homo Sapiens (Human)"
    #organism_name = GetProteinAnnontate(protein_info, "organism_name")
  }
  return(organism_name)
}
```


```{r}
covid_interactions_table = binary_interactions_covid_dedupe_int_dist %>% mutate(
  Species_A = map(Protein_A, getOrganism), 
  Species_B = map(Protein_B, getOrganism)
)
```

covid proteome:
```{r}
covid_proteome = covid_2697049[covid_2697049$Reviewed == "reviewed"]$Entry
```

```{r}
covid_interactions_table %>% filter(Species_B == "Severe acute respiratory syndrome coronavirus 2 (2019-nCoV) (SARS-CoV-2)")
```
There are 5 covid-covid interactions

```{r}
covid_interactions_table_pubs = covid_interactions_table %>% mutate(
  PMIDs = map2_chr(Protein_A, Protein_B, getPMIDs)
)
```


```{r}
getPMIDs = function(proteinA, proteinB) {
  interacts_with = covid_2697049[(covid_2697049$Entry == proteinA),]$`Interacts with` %>% strsplit("; ") %>% unlist()
  if(proteinB %in% interacts_with) {
    pmids = covid_2697049[(covid_2697049$Entry == proteinA),]$`PubMed ID`
  } else {
    pmids = NA
  }
  return(pmids)
}
```


```{r}
fwrite(covid_interactions_table_pubs, file = "/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/DATABASE_PIPELINE/benchmark2/2697049_interactions_table.tsv", sep = "\t")
```


