---
title: "covid papers"
output: html_document
date: "2024-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(data.table)

```

Benchmark 2: 2019-nCoV
Taxonomy ID: 2697049

```{r}
covid_2697049 = fread("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/DATABASE_PIPELINE/benchmark2/uniprotkb_2697049_2024_02_05.tsv")
```

```{r}
table(covid_2697049$Reviewed)
```
Onlyl 17 reviewed cases - will incorporate all unreviewed + reviewed cases for this benchmark
due to the lack of reviewed cases. Ideally, we would have liked to go with just the reviewed cases.

```{r}
length(unique(covid_2697049$`PubMed ID`))
```
101 unique papers that mention covid proteins
note: not quite, they need to be separated and counted uniquely

```{r}
covid_papers = covid_2697049$`PubMed ID` %>% strsplit("; ") %>% compact %>% unlist() %>% unique()
```


```{r}
#setwd("/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/DATABASE_PIPELINE/benchmark2/")
```



```{r}
table(covid_2697049$`Interacts with`)
```
about ~270 binary interactions

```{r}
library(easyPubMed)
```

Resources:
https://www.data-pulse.com/dev_site/easypubmed/
https://www.data-pulse.com/projects/Rlibs/vignettes/easyPubMed_demo.html#Tutorial
"Other than its web portal, PubMed can be programmatically queried via the NCBI Entrez E-utilities interface."
"easyPubMed is an open-source R interface to the Entrez Programming Utilities aimed at allowing programmatic access to PubMed in the R environment. The package is suitable for downloading large number of records, and includes a collection of functions to perform basic processing of the Entrez/PubMed query responses."

easyPubMed pipeline is a 3-step process:
1: query step: get_pubmed_ids()
2: download pubmed records step: fetch_pubmed_data()
3: process/extract data step: depends what you're trying to accomplish

Documentation:
https://cran.r-project.org/web/packages/easyPubMed/easyPubMed.pdf


```{r}
test_pmid_covid = 32511352
test_query_covid = get_pubmed_ids(test_pmid_covid)
test_paper_covid = fetch_pubmed_data(test_query_covid, format = "xml", encoding = "ASCII")
```

```{r}
test_out = batch_pubmed_download(pubmed_query_string = test_pmid_covid,
                                 dest_dir = "/Users/kimg044/Library/CloudStorage/OneDrive-PNNL/Documents/PPI/DATABASE_PIPELINE/benchmark2/test_easyPubMed/",
                                 dest_file_prefix = toString(test_query_covid$IdList$Id),
                                 format = "xml",
                                 batch_size = 1,
                                 encoding = "UTF-8")
```

format "medline" did not include the entire article...
uilist was just the PMID in a text file
abstract was similar output to medline - not entire article
asn.1 was blank?
xml = gross, full of html tags, impossible to read in this format, and perhaps might not include the whole paper either?

https://cran.r-project.org/web/packages/easyPubMed/vignettes/getting_started_with_easyPubMed.html
https://www.data-pulse.com/projects/Rlibs/vignettes/easyPubMed_01_getting_started.html

demo 3: parallelization with doParallel and foreach 
https://www.data-pulse.com/projects/Rlibs/vignettes/easyPubMed_02_advanced_tutorial.html
https://www.data-pulse.com/dev_site/easypubmed/
under the hood,ish-
https://www.data-pulse.com/projects/Rlibs/vignettes/easyPubMed_demo.html#Tutorial
example data:
https://www.data-pulse.com/projects/Rlibs/vignettes/building_the_easyPubMed_EPMsamples_dataset.html
basic explanation:
http://www.biotechworld.it/bioinf/2016/01/21/scraping-pubmed-data-via-easypubmed-xml-and-regex-in-r-for-a-targeting-campaign/
under the hood, github:
https://github.com/dami82/easyPubMed

No package for retrieving full-text from Pubmed, instead PubMed Central:
https://www.biostars.org/p/475309/
https://www.ncbi.nlm.nih.gov/pmc/tools/get-full-text/


